---
title: "Initial Korean covid 19"
author: "Cynthia Cornejo"
date: "6/21/23"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
  
  
# Your Question - Is dying independent of sex?

## METHODS

We shall examine the patients' sex to determine whether they are still alive or have passed away.  I'll make use of the Cat~ Cat variables I have in my collection. This is the kind of issue that we may put into our project from the helper file and use to obtain the solution.


## Graphical results - use ggplot commands to do this  

#### Plot barplots ggplot way

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

I can see from the counting chart that there weren't many deaths. It is 50-70. Males make a roughly 40% of the population, with females making up 50%. Males are less numerous than females. Male fatalities are twice as high. 
  
## Numerical results


## Inferential results 




## Conclusions

