---
title: "Initial Korean covid 19"
author: "Cynthia Cornejo"
date: "6/21/23"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
  
  
# Your Question - Is dying independent of sex?

## METHODS

We shall examine the patients' sex to determine whether they are still alive or have passed away.  I'll make use of the Cat~ Cat variables I have in my collection. This is the kind of issue that we may put into our project from the helper file and use to obtain the solution.


## Graphical results - use ggplot commands to do this  

#### Plot barplots ggplot way

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

I can see from the counting chart that there weren't many deaths. It is 50-70. Males make a roughly 40% of the population, with females making up 50%. Males are less numerous than females. Male fatalities are twice as high. 
  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

I can tell from the percentages that there are 12% more females alive. 36% of the deceased are female and 97% of men are still alive, according to the second graph. Male deaths made up 3% of the total, while female deaths were 1.4%.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

1411 Males and 1831 females out of a group were still alive. Overall, there are more living females than males. In the Chi-Square, the P value is one out of a thousand. Males and females have a very minimal risk of dying. Nine out of 10,000 results in Fisher's test have almost the same P value. Additionally, you can view the odds ratio. The likelihood of females living compared to males passing away. We are interested in the likelihood that this will occur. 

## Conclusions

I discovered how to use the helper file to create a ggplot in today's lecture. I switched the projects variable "kRdata" and the class variable to result. A numerical graph in percent and counting was also something I could do. O also conducted chi- square and Fisher's tests, and the P value for both are the same. The end effect is that we are interested in the likelihood that this will occur. Overall, males had more suffering during the study. Compared to men, women had the advantage. 
